---
title: "Computing missing data by borrowing strength from similar countries"
author: Karen Cao
output: html_notebook
---
```{r, warning = FALSE, include = FALSE, message = FALSE}
# Load all required dependencies
library(openxlsx)
library(tidyverse)
library(lubridate)
library(stats)
library(data.table)
library(gridExtra)
library(ggplot2)
library(FNN)
library(dplyr)
library(broom)
library(tidyr)
library(mosaic)
```

```{r, include = FALSE}
# Load in the data
tfr = read.csv("tfr_data.csv", header = T)
subregion = read.csv("un_subregion.csv", header = T)
tfr = merge(x = tfr, y = subregion[ , -1], by = "ISO.code")
tfr$TimeMid = as.Date.character(format(date_decimal(tfr$TimeMid), "%Y-%m-%d"))
tfr <- tfr %>% mutate(year = year(TimeMid)) %>% filter(year <= 2016)

# Define tfr_m, which takes mean tfr value at each year(when data is available) regardless of reliability of different sources
tfr_m = tfr %>% group_by(year, ISO.code, Country.or.area) %>% summarise(DataValue = mean(DataValue)) %>% arrange(ISO.code, year) %>% ungroup()

# Load in the UN estimates file
un.est = read.csv("wpp2017-fertility-estimates.csv", header = T) %>% filter(Year <= 2016, Indicator == "TFR")
country.index = tfr_m[,2:3] %>% group_by(ISO.code, Country.or.area) %>% summarise(n = n())
country.index = country.index[-3]
un.est = left_join(x = un.est, y = country.index, by = c("LocID" = "ISO.code"))
```

```{r, include = FALSE}
# Define the polynomial function
best_model_na <- function(country, df = tfr_na, tfr_var = "DataValue", max_degree = 9, print.plot = TRUE) {
  ## Create country data frame
  country_data <- df %>%
    filter(Country.or.area == country, year >= 1950, year <= 2016)
  mean_year <- mean(country_data$year)
  country_data <- country_data %>%
    mutate(year_ctr = year - mean_year)
  
  ## Initialize values for loop
  n <- nrow(country_data)
  current_formula <- paste0(tfr_var,"~year_ctr")
  best_model <- lm(formula(current_formula), data = country_data)
  best_degree <- 1
  p_value <- NA
  
  ## Iteratively increase degree of polynomial and test if last degree is significant
  for(i in 2:max_degree) {
    ## Stitch new term onto polynomial
    new_formula <- paste0(current_formula,"+I(year_ctr^",i,")")
    new_model <- lm(formula(new_formula), data = country_data)
    compare_anova <- anova(new_model, best_model)
    
    ## Is it significant? 
    if((compare_anova$`Pr(>F)`[2] < 0.01) & is.na(compare_anova$`Pr(>F)`[2]) == FALSE) {
      best_model <- new_model
      best_degree <- i # If so, save the degree as the best one
      p_value <- compare_anova$`Pr(>F)`[2]
    }
    ## Try the next polynomial degree
    current_formula <- new_formula
  }
  
  year1 <- data.frame(year = 1950:2016, year_ctr = 1950:2016 - mean_year)
  predictions <- predict(best_model, newdata = year1, interval = "confidence")
  country.pred <- tibble(year = 1950:2016) %>%
    mutate(fit = predictions[,1],
           lwr = predictions[,2],
           upr = predictions[,3])
  
  p <- ggplot() + 
      geom_point(data = country_data, aes(year, DataValue, colour = factor(Ori))) + 
      geom_line(data = country.pred, aes(year, fit)) + 
      geom_line(data = un.est %>% filter(Country.or.area == country), aes(Year, DataValue, color = "red")) +
      geom_ribbon(data = country.pred, aes(year, ymin = lwr, ymax = upr), alpha = 0.4) + 
      scale_y_continuous(limits = c(0, 11.5)) + 
      ggtitle(paste(country, ", degree = ", best_degree)) +
      lims(x = c(1950,2016)) + 
      scale_colour_discrete(name  ="Imputed?", breaks = c(0, 1), labels = c("Yes", "No"))
    
  ## Print plot if specified
  if(print.plot) {
    print(p)
  }
  
  ## Create the object to be returned
  return_object <- list(model = best_model)
  return_object$data <- country_data
  return_object$details <- list(degree = best_degree, 
                                      R2 = summary(best_model$adj.r.squared),
                                      RMSE = sqrt(sum(residuals(best_model)^2)/length(residuals(best_model))))
  return_object$predictions <- country.pred
  return_object$p.value <- p_value
  return_object$plot <- p
  
  return(return_object)
}
```

```{r, include = FALSE}
# Create scaled covariates for ALL countries spanning from 1950 to 2016
all_covariates <- read.csv("wpp2017-covariates.csv", stringsAsFactors = FALSE) %>% filter(Year >= 1950, Year <= 2016)
all_covariates = spread(all_covariates, Indicator, DataValue)
covariates = left_join(x= all_covariates, y = country.index, by = c("LocID" = "ISO.code"))
covariates[, 3:13] = apply(covariates[, 3:13], 2, zscore) # scale all values
write.csv(covariates, file = "wpp_covariates_fulltime.csv")
```

# PCA
```{r, warning = FALSE, message = FALSE}
# Read in the dataset that contains scaled covariates for all countries
all_covariates = read.csv("wpp_covariates_fulltime.csv", stringsAsFactors = FALSE) %>% filter(LocID %in% unique(tfr_m$ISO.code))

# Perform PCA
pca_post = prcomp(~ PopDensity + PopGrowth + PropUrban + Q1535B + Q1535F + Q1535M + Q1545B + Q1545F + Q1545M + TotalPop + U5MR, data = all_covariates)
covariates_pcs_post <- bind_cols(all_covariates %>% select(Country.or.area), as.data.frame(pca_post$x)) %>% group_by(Country.or.area) %>% summarise_at(2:12, mean) %>% ungroup()
```

# For each country, find its 3 most similar countries using knn
```{r}
# summary(pca_post) 

# Find the 3 most similar countries using the first 6 PCs
# The first 6 PCs explain 99% of the variance
knn = knn.index(dist(covariates_pcs_post %>% select(PC1:PC6)), k = 3)
knn.name = as.data.frame(knn)
for (i in 1:201) {
  knn.name[i, ] <- covariates_pcs_post$Country.or.area[knn[i, ]]
}
rownames(knn.name) <- covariates_pcs_post$Country.or.area

# knn.name is the data frame that lists the three most similar countries for every single country

head(knn.name)
```

# Compue missing data
Wheneve the tfr data value is missing for a country at a specific year, this missing data is calculated by taking the averaged of the tfr value that year from its 3 most similar countries. The "final" dataset contains year, name of the country, tfr value and a binary indicator indicator whether the tfr value that year is original or imputed (i.e. bowrrowed strength from its three most similar countries).

```{r}
tfr_fillna = list()

for(i in 1:201){
  tfr_L = tfr_m %>% filter(Country.or.area == unique(tfr_m$Country.or.area)[i])
  if(unique(tfr_m$Country.or.area)[i] %in% rownames(knn.name)){
    tfr_R = tfr_m %>% filter(Country.or.area %in% knn.name[unique(tfr_m$Country.or.area)[i], ]) %>% arrange(year) %>% group_by(year) %>% summarise(DataValue = mean(DataValue)) %>% ungroup()
    tfr_R$Country.or.area = rep(unique(tfr_m$Country.or.area)[i], nrow(tfr_R))
    tfr_fill = full_join(tfr_L, tfr_R, by = "year") %>% arrange(year)
    tfr_fill$Country.or.area = ifelse(is.na(tfr_fill$Country.or.area.x), as.character(tfr_fill$Country.or.area.y), as.character(tfr_fill$Country.or.area.x))
    tfr_fill$DataValue = ifelse(is.na(tfr_fill$DataValue.x), tfr_fill$DataValue.y, tfr_fill$DataValue.x)
    tfr_fill$Ori = ifelse(is.na(tfr_fill$DataValue.x), 0, 1) # indicate whether DataValue is borrowed from other countries/areas
    var_keep = c("year", "Country.or.area", "DataValue", "Ori")
    tfr_fillna[[i]] = tfr_fill[, var_keep]
  }else{
    tfr_fillna[[i]] = tfr_L
  }
}

tfr_na = dplyr::bind_rows(tfr_fillna)
head(tfr_na)
```

# Plots
Look at plots for all countries by fitting the baseline (polynomial) model. All plots are saved in a pdf file called "Compute missing valye using knn".
```{r, include = FALSE, warning = FALSE, message = FALSE}
# plot_list = list()
# plot_degree = vector()
# country.parse = unique(tfr_na$Country.or.area)

# for(i in 1:length(country.parse)) {
#    country <- country.parse[i]
#    plot_list[[i]] <- best_model_na(country, print.plot = FALSE)$plot
#    plot_degree[i] <- best_model_na(country, print.plot = FALSE)$details$degree # create degree of the polynomial model for each country
#    if(i %% 10 == 0) {
#      print(paste("Country:", country))
#    }
# }

# ggsave("Compute missing value using knn.pdf", marrangeGrob(grobs = plot_list, nrow = 3, ncol = 2))
```

Look at a few selected plots. 
```{r}
# Good ones
best_model_na("Angola", print.plot = FALSE)$plot
best_model_na("Serbia", print.plot = FALSE)$plot

# Bad ones
best_model_na("Mayotte", print.plot = FALSE)$plot
best_model_na("Saudi Arabia", print.plot = FALSE)$plot
```

And look at the 3 most similar countries for these 4 countries
```{r}
countr.list = c("Angola", "Serbia", "Mayotte", "Saudi Arabia")
knn.name.select = knn.name %>% filter(rownames(knn.name) %in% countr.list)
rownames(knn.name.select) = countr.list
colnames(knn.name.select) = c("1", "2", "3")
knn.name.select
```

