---
title: "PolyBaselineModel"
author: "Karen Cao"
date: "11/15/2018"
output: pdf_document
---

# Polynomial Baseline Model

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(openxlsx)
library(tidyverse)
library(lubridate)
library(stats)
library(data.table)
library(dplyr)
```

## Data Pre-Processiong

* merge tfr data with subregion data
* calculate mean tfr value for each year, ignore reliability of different sources for now
* limit the data only from 1950 to 2016

```{r, include = FALSE}
tfr = read.csv("tfr_data.csv", header = T)
subregion = read.csv("un_subregion.csv", header = T)
tfr = merge(x = tfr, y = subregion[ , -1], by = "ISO.code")
tfr$TimeMid = as.Date.character(format(date_decimal(tfr$TimeMid), "%Y-%m-%d"))
tfr_m = tfr %>% group_by(year = year(TimeMid), ISO.code, Country.or.area) %>% summarise(DataValue = mean(DataValue)) %>% arrange(ISO.code, year) %>% ungroup() %>% filter(year >= 1950, year <= 2016)
```

## Polynomial Model

* time trend of tfr over year
* limit to max degree of 9
* center year to avoid multicollinearity
* add higher-ordered term till variance explained is no longer significant, compare with anova
* use significant level of 0.05

```{r, include = FALSE}
best_model <- function(country, max_degree = 9, print.plot = TRUE) {
  country_data <- tfr_m %>%
    filter(Country.or.area == country)
  mean_year <- mean(country_data$year)
  country_data <- country_data %>%
    mutate(year_ctr = year - mean_year)
  n <- nrow(country_data)
  current_formula <- "DataValue~year_ctr"
  best_model <- lm(formula(current_formula), data = country_data)
  best_degree <- 1
  p_value <- NA
  
  for(i in 2:max_degree) {
    new_formula <- paste0(current_formula,"+I(year_ctr^",i,")")
    new_model <- lm(formula(new_formula), data = country_data)
    compare_anova <- anova(new_model, best_model)
    if((compare_anova$`Pr(>F)`[2] < 0.05) & is.na(compare_anova$`Pr(>F)`[2]) == FALSE) {
      best_model <- new_model
      best_degree <- i
      p_value <- compare_anova$`Pr(>F)`[2]
    } else {
      NULL
    }
    current_formula <- new_formula
  }
  
  year1 <- unique(country_data$year)
  country.pred <- data.frame(year = year1, predict(best_model, interval = 'confidence'))
  
  p <- ggplot() + 
    geom_point(data = country_data, aes(year, DataValue)) + 
    geom_line(data = country.pred, aes(year, fit, color = "red")) + 
    geom_ribbon(data = country.pred, aes(year, ymin = lwr, ymax = upr), alpha = 0.4) + 
    ggtitle(paste(country, ", degree = ", best_degree)) + 
    theme(legend.position="none") + 
    ylim(0, 12)
  
  return(list(
    model_deets = list(degree = best_degree,
                       R2 = summary(best_model$adj.r.squared),
                       RMSE = sqrt(sum(residuals(best_model)^2)/length(residuals(best_model)))
    ),
    model = best_model, p.value = p_value,
    plot = p)
  )
}
```

```{r, include = FALSE}
plot_list <- vector("list", n_distinct(tfr_m$Country.or.area))
plot_degree <- rep(NA, n_distinct(tfr_m$Country.or.area))

for(i in 1:n_distinct(tfr_m$Country.or.area)) {
  countries <- sort(unique(tfr_m$Country.or.area))
  country <- countries[i]
  plot_list[[i]] <- best_model(country, print.plot = FALSE)$plot
  plot_degree[i] <- best_model(country, print.plot = FALSE)$model_deets$degree # create degree of the polynomial model for each country
  if(i %% 10 == 0) {
    print(paste("Country:", country))
  }
}
```

```{r}
# order the plots by degrees, from the highest to the lowest
polyplot = data.frame(index = 1:201, country.or.area = unique(tfr_m$Country.or.area), degree = plot_degree) %>% arrange(desc(degree))
plot_list_ordered <- vector("list", n_distinct(tfr_m$Country.or.area))

for(i in 1:201){
  plot_list_ordered[[i]] = plot_list[[polyplot$index[i]]]
}
```

```{r, warning=FALSE}
ggsave("BaselineModelPlot.pdf", marrangeGrob(grobs = plot_list_ordered, nrow = 3, ncol = 2))
```

